---
title: "German Credit Data - Risk Analysis using Logistic Regression"
author: "Anumeha Dwivedi"
date: "4 February 2018"
output: rmarkdown::github_document
---

###Loading libraries
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(Hmisc)
library(corrplot)
library(leaps)
library(glmnet)
library(ROCR)
library(verification)
library(psych)
```

###Exploratory Data Analysis (EDA) and Data Pre-processing
```{r, message=FALSE, warning=FALSE}
cc_data <- read_csv("/Users/tinycloud/Documents/Spring 18/Data Mining 1/Data/German_Credit_Card_Data.csv")
num_var <- c(2,5,8,11,13,16,18)
factor_var <- c(1,3,4,6,7,9,10,12,14,15,17,19,20,21)
cc_data[,factor_var] <- lapply(cc_data[,factor_var] , factor)
cc_data[num_var] <- sapply(cc_data[num_var] , as.numeric)
summary(cc_data[factor_var])
```

```{r, fig.height=10, fig.width=6, message=FALSE, warning=FALSE}
par(mfrow = c(4,2))
hist(cc_data$Duration, freq = FALSE)
lines(density(cc_data$Duration), col= "blue")
hist(cc_data$`Credit Amount`, probability = TRUE)
lines(density(cc_data$`Credit Amount`), col= "blue")
hist(cc_data$Installmant_rate, probability = TRUE)
lines(density(cc_data$Installmant_rate), col= "blue")
hist(cc_data$Present_residence, probability = TRUE)
lines(density(cc_data$Present_residence), col= "blue")
hist(cc_data$Age, probability = TRUE)
lines(density(cc_data$Age), col= "blue")
hist(cc_data$`No of credits`, probability = TRUE)
lines(density(cc_data$`No of credits`), col= "blue")
hist(cc_data$`Liable people`, probability = TRUE)
lines(density(cc_data$`Liable people`), col= "blue")

par(mfrow = c(4,4))
barplot(table(cc_data$Existing_checking_account))
barplot(table(cc_data$Credit_history))
barplot(table(cc_data$Purpose))
barplot(table(cc_data$`Savings account`))
barplot(table(cc_data$`Present employment since`))
barplot(table(cc_data$`Personal status and Tax`))
barplot(table(cc_data$`Other debtors / guarantors`))
barplot(table(cc_data$Property))
barplot(table(cc_data$`Other installment plans`))
barplot(table(cc_data$Housing))
barplot(table(cc_data$Job))
barplot(table(cc_data$Telephone))
barplot(table(cc_data$`Foreign Worker`))
barplot(table(cc_data$Default))
```

###Splitting into test and train
```{r, message=FALSE, warning=FALSE}
set.seed(12363925)
index <- sample(1:nrow(cc_data), 0.8*nrow(cc_data))
train <- cc_data[index,]
test <- cc_data[-index,]
```

###Trying different link functions
```{r, message=FALSE, warning=FALSE, results='hide'}
log.fit <- glm(Default ~ Existing_checking_account + Duration + Credit_history + Purpose + 
                  `Credit Amount` + `Savings account` + Installmant_rate + `Other debtors / guarantors` 
                + `Other installment plans` + `No of credits` + Telephone + `Foreign Worker`, 
                data = train, family = binomial)

log.fit.probit <- glm(Default ~ Existing_checking_account + Duration + Credit_history + Purpose + 
                 `Credit Amount` + `Savings account` + Installmant_rate + `Other debtors / guarantors` 
               + `Other installment plans` + `No of credits` + Telephone + `Foreign Worker`, 
               data = train, family = binomial(link = "probit"))

log.fit.cloglog <- glm(Default ~ Existing_checking_account + Duration + Credit_history + Purpose + 
                 `Credit Amount` + `Savings account` + Installmant_rate + `Other debtors / guarantors` 
               + `Other installment plans` + `No of credits` + Telephone + `Foreign Worker`, 
               data = train, family = binomial(link = "cloglog"))

summary(log.fit)
summary(log.fit.probit)
summary(log.fit.cloglog)
```

####Comparing AIC and BIC for the models with different links
```{r, message=FALSE, warning=FALSE}
AIC(log.fit)
AIC(log.fit.probit)
AIC(log.fit.cloglog)

BIC(log.fit)
BIC(log.fit.probit)
BIC(log.fit.cloglog)
```

###Hit rates and ROC curves for the 3 links on training data
```{r, message=FALSE, warning=FALSE}
log_train_pred <- predict(log.fit, type = "response")
prob_train_log <- ifelse(log_train_pred > 0.5, 2, 1)
mean(prob_train_log == train$Default) 
probit_train_pred <- predict(log.fit.probit, type = "response")
prob_train_probit <- ifelse(probit_train_pred > 0.5, 2, 1)
mean(prob_train_probit == train$Default)
cloglog_train_pred <- predict(log.fit.cloglog, type = "response")
prob_train_cloglog <- ifelse(cloglog_train_pred > 0.5, 2, 1)
mean(prob_train_cloglog == train$Default)

roc.plot(x = train$Default == "2", pred = cbind(log_train_pred, probit_train_pred,cloglog_train_pred), 
         legend = TRUE, leg.text = c("logit", "Probit", "cloglog"))$roc.vol
```

###Hit rates and ROC curves for the 3 links on test data
```{r, message=FALSE, warning=FALSE}
log_test_pred <- predict(log.fit, newdata = test, type = "response")
prob_test_log <- ifelse(log_train_pred > 0.5, 2, 1)
mean(prob_test_log == test$Default)
probit_test_pred <- predict(log.fit.probit, newdata = test, type = "response")
prob_test_probit <- ifelse(probit_train_pred > 0.5, 2, 1)
mean(prob_test_probit == test$Default)
cloglog_test_pred <- predict(log.fit.cloglog, newdata = test, type = "response")
prob_test_cloglog <- ifelse(cloglog_train_pred > 0.5, 2, 1)
mean(prob_test_cloglog == test$Default)

roc.plot(x = test$Default == "2", pred = cbind(log_test_pred, probit_test_pred,cloglog_test_pred), 
         legend = TRUE, leg.text = c("logit", "Probit", "cloglog"))$roc.vol
```

###Variable selection - step AIC, step BIC and LASSO
```{r, message=FALSE, warning=FALSE, results='hide'}
null.model <- glm(Default ~ 1, data = train, family = "binomial")
logstepAIC <- step(null.model, scope = list(lower=null.model, upper=log.fit), k = 2, direction="both")
logstepBIC <- step(null.model, scope = list(lower=null.model, upper=log.fit), direction = "both", 
                   k = log(nrow(train)))

##Lasso
cc_data_scale <- cc_data
cc_data_scale[num_var] <- scale(cc_data_scale[num_var])
cc_data_scale[factor_var] <- sapply(cc_data[factor_var] , as.numeric)
X.train<- as.matrix(cc_data_scale)[index,1:20]
X.test<-  as.matrix(cc_data_scale)[-index,1:20]
Y.train <- cc_data_scale$Default[index]
Y.test <- cc_data_scale$Default[-index]
lasso_fit <- glmnet(x = X.train ,y = Y.train ,alpha = 1 , family = "binomial")
cv.out <- cv.glmnet(x = X.train ,y = Y.train ,alpha = 1 , family = "binomial" ,type.measure = "mse")
plot(cv.out, xvar = "lambda", label=TRUE)
#min value of lambda
lambda_min <- cv.out$lambda.min
#best value of lambda
lambda_1se <- cv.out$lambda.1se
#regression coefficients
coef(cv.out,s=lambda_1se)

logstepAIC$anova
logstepBIC$anova
coef(cv.out,s=lambda_1se)

AIC(logstepAIC)
AIC(logstepBIC)

BIC(logstepAIC)
BIC(logstepBIC)
```

###Model selection
```{r, message=FALSE, warning=FALSE}
stepAICpred <- predict(logstepAIC, type = "response")
prob_train_AIC <- ifelse(stepAICpred > 0.5, 2, 1)
mean(prob_train_AIC == train$Default)
stepBICpred <- predict(logstepBIC, type = "response")
prob_train_BIC <- ifelse(stepBICpred > 0.5, 2, 1)
mean(prob_train_BIC == train$Default)
lasso_prob_train <- predict(cv.out, newx = X.train, s = lambda_1se, type = "response")
prob_train_lasso <- ifelse(lasso_prob_train > 0.5, 2, 1)
mean(prob_train_lasso == train$Default)

roc.plot(x = train$Default == "2", pred = cbind(stepAICpred, stepBICpred, lasso_prob_train), 
         legend = TRUE, leg.text = c("Step AIC", "Step BIC", "LASSO"))$roc.vol

stepAICpred <- predict(logstepAIC, newdata = test, type = "response")
prob_test_AIC <- ifelse(stepAICpred > 0.5, 2, 1)
mean(prob_test_AIC == test$Default)
stepBICpred <- predict(logstepBIC, newdata = test, type = "response")
prob_test_BIC <- ifelse(stepBICpred > 0.5, 2, 1)
mean(prob_test_BIC == test$Default)
lasso_prob_test <- predict(cv.out, newx = X.test, s = lambda_1se, newdata = test, type = "response")
prob_test_lasso <- ifelse(lasso_prob_test > 0.5, 2, 1)
mean(prob_test_lasso == test$Default)

roc.plot(x = test$Default == "2", pred = cbind(stepAICpred, stepBICpred, lasso_prob_test), 
         legend = TRUE, leg.text = c("Step AIC", "Step BIC", "LASSO"))$roc.vol
```

###Finding Optimal Probability
```{r, message=FALSE, warning=FALSE}
logstepAIC

searchgrid = seq(0.01, 0.99, 0.01)
result = cbind(searchgrid, NA)

# in the cost function, both r and pi are vectors, r=truth, pi=predicted probability
cost1 <- function(r, pi) 
{
  weight1 = 5
  weight0 = 1
  c1 = (r == 2) & (pi < pcut)  #logical vector - true if actual 1 but predict 0
  c0 = (r == 1) & (pi > pcut)  #logical vecotr - true if actual 0 but predict 1
  return(mean(weight1 * c1 + weight0 * c0))
}

for (i in 1:length(searchgrid)) 
{
  pcut <- result[i, 1]
  # assign the cost to the 2nd col
  result[i, 2] <- cost1(train$Default, predict(logstepAIC, type = "response"))
}

plot(result, ylab = "Cost in Training Set", col = "red")
result[which(result[,2] == min(result[,2])),]

final_model <- glm(Default ~ Existing_checking_account + Duration + Credit_history + Purpose + 
                         `Credit Amount` + `Savings account` + Installmant_rate + `Other debtors / guarantors` 
                       + `Other installment plans` + `No of credits` + Telephone + `Foreign Worker`, 
                       data = cc_data, family = binomial)
```
