---
title: "Using KNN for Classification"
author: "Anumeha Dwivedi"
date: "10 January 2018"
output: rmarkdown::github_document
---

###Loading Libraries
```{r Loading Libraries, message=FALSE, warning=FALSE}
library(class)
library(ggplot2)
library(gmodels)
library(GGally)
```


###Understanding Data
```{r Understanding Data, message=FALSE, warning=FALSE}
str(iris)
summary(iris)
ggpairs(iris)
```


###Data Visualization
```{r Visualization, message=FALSE, warning=FALSE}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point()

ggplot(data = iris, aes(x = Petal.Length, y = Petal.Width, col = Species)) +
  geom_point()
```


###Scaling the data
```{r Scaling, message=FALSE, warning=FALSE}
scale_iris <- iris
scale_iris[,1:4] <- scale(iris[,1:4])
```


###Splitting training and test sets
```{r, message=FALSE, warning=FALSE}
set.seed(123)
index <- sample(1:nrow(scale_iris), nrow(scale_iris)*0.6)
iris_train <- scale_iris[index,]
iris_test <- scale_iris[-index,]
```

###KNN Model
```{r Applying KNN for multiple values of K, message=FALSE, warning=FALSE}
error <- c()
for (i in 1:15)
{
  knn.fit <- knn(train = iris_train[,1:4], test = iris_test[,1:4], cl = as.factor(iris_train$Species), k = i)
  error[i] = 1- mean(knn.fit == iris_test$Species)
}

ggplot(data = data.frame(error), aes(x = 1:15, y = error)) +
  geom_line(color = "Blue")
```

We see that for K = 5, we have the smallest error 

###Checking accuracy of the final model
```{r Checking Accuracy, message=FALSE, warning=FALSE}
knn.fit <- knn(train = iris_train[,1:4], test = iris_test[,1:4], cl = as.factor(iris_train$Species), k = 5)

CrossTable(x = iris_test$Species, y = knn.fit,prop.chisq=FALSE) 
```

From the above matrix, we see that all our classification for Setosa is correct, we have classified 1 Virginica as Versicolor and 2 Versicolor as Virginica. Overall the accurcy of our model is 57/60 = 95%.